---
title: "Introduction to R"
author: "Christine Choirat"
date: "March 27, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to R

## Links of interest

https://cran.r-project.org/

https://www.rstudio.com/

https://www.rstudio.com/products/rstudio/#Desktop

http://dss.iq.harvard.edu/workshop-materials

http://tutorials.iq.harvard.edu/R/Rintro/Rintro.html

# R on Odyssey

## Installing R packages

### File `.bash.rc`

```{bash eval=FALSE}
# User specific aliases and functions
source new-modules.sh

module load R

export R_LIBS_USER=/n/[...]/[...]/apps/R:$R_LIBS_USER
```

```{r eval=FALSE}
install.packages('package_name', dependencies=TRUE, 
                 repos='http://cran.rstudio.com/')
```

For example:

```{r eval=FALSE}
install.packages('tidyverse', dependencies=TRUE, 
                 repos='http://cran.rstudio.com/')
```

## Using `devtools`

`install_bioc`

`install_cran`

`install_github`

`install_url`

`install_version`

## Going parallel locally

```{r eval=FALSE}
library(parallel)
detectCores() # 4
```

```{r eval=FALSE}
meanDist <- function(n, nsamp = 500) {
  replicate(nsamp, mean(rnorm(n)))
}

## range of sample sizes to iterate over
sampSizes <- seq(10, 500, by = 5)
system.time(means <- lapply(sampSizes, meanDist))
```

```{txt}
    user  system elapsed 
   1.450   0.031  17.016
```

## Going parallel locally

```{r eval=FALSE}
library(parallel) 
system.time(means <- mclapply(sampSizes, meanDist, mc.cores = 2))
```

```{txt}
    user  system elapsed 
    0.736   0.048   7.592 
```

## Going parallel on Odyssey: interactive jobs

```{bash eval=FALSE}
srun -p interact --pty --mem 500 -t 0-10:00 /bin/bash
```

```{bash eval=FALSE}
[cchoirat@rclogin04 ~]$ pwd
/n/home03/cchoirat
[cchoirat@rclogin04 ~]$ srun -p interact --pty --mem 500 -t 0-10:00 /bin/bash
[cchoirat@holy2a18308 ~]$ R
```

## Going parallel on Odyssey

```{txt}
> library(parallel) 
> detectCores()
[1] 64
> meanDist <- function(n, nsamp = 500) {
+   replicate(nsamp, mean(rnorm(n)))
+ }
> 
> sampSizes <- seq(10, 500, by = 5)
> system.time(means <- lapply(sampSizes, meanDist))
   user  system elapsed 
  5.038   0.010   5.048 
> library(parallel) 
> system.time(means <- mclapply(sampSizes, meanDist, mc.cores = 2))
   user  system elapsed 
  2.521   0.040   5.162 
> system.time(means <- mclapply(sampSizes, meanDist, mc.cores = 4))
   user  system elapsed 
  6.381   0.138   5.258 
```

## Going parallel on Odyssey

```{txt}
> system.time(means <- mclapply(sampSizes, meanDist, mc.cores = 32))
srun: Job step aborted: Waiting up to 32 seconds for job step to finish.
srun: error: holy2a18308: task 0: Killed
```

## Going parallel on Odyssey: SLURM and job arrays

https://rc.fas.harvard.edu/resources/documentation/submitting-large-numbers-of-jobs-to-odyssey/

http://airoldilab.github.io/odyssey/

`%A` in the `#SBATCH` line becomes the job ID

`%a` in the `#SBATCH` line becomes the array index

`${SLURM_ARRAY_TASK_ID}` is a shell variable that is set when
the job runs,  and it is substituted into the parameter to 
generate the proper filename which is passed on to tophat
